### 2018-03-01 submission/manuscript.Rmd
update with new orders after pats edits

***Combined taxa model classifies subjects better than using individual taxa:*** 
Instead of attempting to classify subjects based on individual taxa, next we generated Random Forest models that combined the individual taxa and evaluated the ability to classify as subject’s disease status. For data from stool samples, the combined model had an AUC of `r round(median(filter(crc_stool_model_aucs, study == train_model, model_type == "full_model")[, "AUC"]), digits = 2)`, which was significantly higher than any of the AUC values for the individual taxa (P-value < `r round(max(stool_ind_vs_full_taxa$pvalue), digits = 3)`). For the full taxa models using stool, *Bacteroides* and *Lachnospiraceae* were the most common taxa in the top 10% mean decrease in accuracy (MDA) across studies. Similarly, using data from the unmatched tissue samples, the combined model had an AUC of `r round(median(filter(crc_tissue_model_aucs, study == train_model, model_type == "full_model", type == "unmatched")[, "AUC"]), digits = 2)`, which was significantly higher than the AUC values for Blautia and Weissella (P-value < `r round(max(filter(unmatched_tissue_ind_vs_full_taxa, taxa != "Dorea")[, "pvalue"]), digits = 3)`).  For the full taxa models using unmatched tissue, *Lachnospiraceae*, *Bacteroidaceae*, and *Ruminococcaceae* were the most common taxa in the top 10% mean decrease in accuracy across studies. Clearly, pooling the information from the taxa with significant ORs results in a model that outperforms classifications made using individual taxa.

***Performance of models based on taxa relative abundance in full community are not significantly better than those based on taxa with significant ORs:*** 
Next, we asked whether a Random Forest classification model built using all of the taxa found in the communities would outperform the models generated using those taxa with a significant OR. Similar to our inability to identify taxa associated with a significant OR for the presence of adenomas, the median AUCs to classify subjects as having normal colons or having adenomas using data from stool or tissue samples were marginally better than 0.5 for any study[Figure 4A & S3A]. In contrast, the models for classifying subjects as having normal colons or having carcinomas using data from stool or tissue samples yielded AUC values meaningfully higher than 0.5 [Figure 4B & S3B-C]. When we compared the models based on all of the taxa in a community to models based on the taxa with significant ORs, the results were mixed. Using the data from stool samples we found that although the AUC for 6 of 7 studies increased (mean decrease = `r round(mean((filter(crc_stool_model_aucs, study == train_model, model_type == "full_model", study != "flemer")[, "AUC"] - filter(crc_stool_model_aucs, study == train_model, model_type == "select_model", study != "flemer")[, "AUC"])*100), digits = 2)`%), the more expansive models performed worse for 1 of the studies (decrease = `r round(abs(mean((filter(crc_stool_model_aucs, study == train_model, model_type == "full_model", study == "flemer")[, "AUC"] - filter(crc_stool_model_aucs, study == train_model, model_type == "select_model", study == "flemer")[, "AUC"])*100)), digits = 2)`%). The overall improvement in performance was statistically significant (one-tailed paired T-test; P-value = `r round(t.test(filter(crc_stool_model_aucs, study == train_model, model_type == "full_model")[, "AUC"], filter(crc_stool_model_aucs, study == train_model, model_type == "select_model")[, "AUC"], alternative = "greater", paired = TRUE)$p.value, digits = 3)`). Of the `r length(crc_OR_stool$measure)` taxa with significant ORs, all `r length(crc_OR_stool$measure)` were among the top 10% most important taxa as measured by mean decrease in accuracy, in at least one study . Similarly, using the data from unmatched tissue samples we found that the AUC for 4 out of 4 studies decreased between full versus select OR models (mean decrease = `r round(mean((filter(crc_tissue_model_aucs, study == train_model, model_type == "full_model", type == "unmatched")[, "AUC"] - filter(crc_tissue_model_aucs, study == train_model, model_type == "select_model", type == "unmatched")[, "AUC"])*100), digits = 2)`%, one-tailed paired T-test; P-value = `r round(t.test(filter(crc_tissue_model_aucs, study == train_model, model_type == "full_model", type == "unmatched")[, "AUC"], filter(crc_tissue_model_aucs, study == train_model, model_type == "select_model", type == "unmatched")[, "AUC"], alternative = "greater", paired = TRUE)$p.value, digits = 3)`). Of the `r length(crc_unmatched_OR_tissue$measure)` taxa with significant ORs, all `r length(crc_unmatched_OR_tissue$measure)` were among the top 10% most important taxa as measured by mean decrease in accuracy, in at least one study. These results were surprising because it demonstrated that the ability to classify subjects could be done based on a limited characterization of the communities.

@@

***Generalizability of taxon-based models trained on one dataset to the other datasets:*** 
Considering the good performance of the Random Forest models using taxa with a significant OR and using all of the taxa, we next asked how well the models would perform when given data from a different subject cohort. For instance, if a model was trained using data from the Ahn study, we wanted to know how well it would perform using the data from the Baxter study. We found the models trained using the taxa with a significant OR all had a higher median AUC than the models trained using all of the taxa when tested on the other datasets [Figure 5]. As might be expected, the difference between the performance of the modelling approaches appeared to vary with the size of the training cohort. These data suggest that given a sufficient number of subjects with normal colons and carcinomas, Random Forest models trained using a small number of taxa can accurately classify individuals from a different cohort.

@@

**Figure 5: Stool Random Forest Genus-Based Model Test AUCs.** A) Test AUCs of adenoma models using all genera across study. B) Test AUCs of carcinoma models using all genera or significant OR taxa only. The black line represents the AUC at 0.5. The red lines represent the median AUC of all test AUCs for a specific study.

**Figure 6: Most Common Taxa Across Carcinoma Full Community Stool Study Models.** A) Common taxa in the top 10 percent for carcinoma Random Forest all taxa-based models. B) Common taxa in the top 10 percent for carcinoma Random Forest all OTU-based models.

@@

**Figure S2: Tissue Random Forest Model Train AUCs.** A) Adenoma random forest model AUCs between all genera, all OTU, and select model based on significant OR taxa in unmatched and matched tissue. B) Carcinoma random forest model AUCs between all genera, all OTU, and select model based on significant OR taxa in unmatched and matched tissue. The black line represents the median AUC for the respective group. If no values are present in the singificant OR taxa group then there were no significant taxa identified and no model was tested.

**Figure S3: Tissue Random Forest Genus-Based Model Test AUCs.** A) Test AUCs of adenoma models using all genera across study. B) Test AUCs of carcinoma models using all genera for matched tissue studies. C) Test AUCs of carcinoma models using all genera or significant OR taxa only for unmatched tissue studies The black line represents the AUC at 0.5. The red lines represent the median AUC of all test AUCs for a specific study.

**Figure S4: Most Common Genera Across Full Community Tissue Study Models.** A) Common genera in the top 10 percent for matched carcinoma Random Forest all genera-based models. B) Common genera in the top 10 percent for unmatched carcinoma Random Forest all genera-based models. C) Common genera in the top 10 percent for matched carcinoma Random Forest all OTU-based models. D) Common genera in the top 10 percent for unmatched carcinoma Random Forest all OTU-based models.

>>>>

***Combined taxa model classifies subjects better than using individual taxa:*** 
Instead of attempting to classify subjects based on individual taxa, next we generated Random Forest models that combined the individual taxa and evaluated the ability to classify as subject’s disease status. For data from stool samples, the combined model had an AUC of `r round(median(filter(crc_stool_model_aucs, study == train_model, model_type == "full_model")[, "AUC"]), digits = 2)`, which was significantly higher than any of the AUC values for the individual taxa (P-value < `r round(max(stool_ind_vs_full_taxa$pvalue), digits = 3)`). For the full taxa models using stool, *Bacteroides* and *Lachnospiraceae* were the most common taxa in the top 10% mean decrease in accuracy (MDA) across studies [Figure S2]. Similarly, using data from the unmatched tissue samples, the combined model had an AUC of `r round(median(filter(crc_tissue_model_aucs, study == train_model, model_type == "full_model", type == "unmatched")[, "AUC"]), digits = 2)`, which was significantly higher than the AUC values for Blautia and Weissella (P-value < `r round(max(filter(unmatched_tissue_ind_vs_full_taxa, taxa != "Dorea")[, "pvalue"]), digits = 3)`).  For the full taxa models using unmatched tissue, *Lachnospiraceae*, *Bacteroidaceae*, and *Ruminococcaceae* were the most common taxa in the top 10% mean decrease in accuracy across studies [Figure S3]. Clearly, pooling the information from the taxa with significant ORs results in a model that outperforms classifications made using individual taxa.

***Performance of models based on taxa relative abundance in full community are better than those based on taxa with significant ORs:*** 
Next, we asked whether a Random Forest classification model built using all of the taxa found in the communities would outperform the models generated using those taxa with a significant OR. Similar to our inability to identify taxa associated with a significant OR for the presence of adenomas, the median AUCs to classify subjects as having normal colons or having adenomas using data from stool or tissue samples were marginally better than 0.5 for any study[Figure 4A & S4A]. In contrast, the models for classifying subjects as having normal colons or having carcinomas using data from stool or tissue samples yielded AUC values meaningfully higher than 0.5 [Figure 4B & S4B-C]. When we compared the models based on all of the taxa in a community to models based on the taxa with significant ORs, the results were mixed. Using the data from stool samples we found that although the AUC for 6 of 7 studies increased (mean decrease = `r round(mean((filter(crc_stool_model_aucs, study == train_model, model_type == "full_model", study != "flemer")[, "AUC"] - filter(crc_stool_model_aucs, study == train_model, model_type == "select_model", study != "flemer")[, "AUC"])*100), digits = 2)`%), the more expansive models performed worse for 1 of the studies (decrease = `r round(abs(mean((filter(crc_stool_model_aucs, study == train_model, model_type == "full_model", study == "flemer")[, "AUC"] - filter(crc_stool_model_aucs, study == train_model, model_type == "select_model", study == "flemer")[, "AUC"])*100)), digits = 2)`%). The overall improvement in performance was statistically significant (one-tailed paired T-test; P-value = `r round(t.test(filter(crc_stool_model_aucs, study == train_model, model_type == "full_model")[, "AUC"], filter(crc_stool_model_aucs, study == train_model, model_type == "select_model")[, "AUC"], alternative = "greater", paired = TRUE)$p.value, digits = 3)`). Of the `r length(crc_OR_stool$measure)` taxa with significant ORs, all `r length(crc_OR_stool$measure)` were among the top 10% most important taxa as measured by mean decrease in accuracy, in at least one study . Similarly, using the data from unmatched tissue samples we found that the AUC for 4 out of 4 studies decreased between full versus select OR models (mean decrease = `r round(mean((filter(crc_tissue_model_aucs, study == train_model, model_type == "full_model", type == "unmatched")[, "AUC"] - filter(crc_tissue_model_aucs, study == train_model, model_type == "select_model", type == "unmatched")[, "AUC"])*100), digits = 2)`%, one-tailed paired T-test; P-value = `r round(t.test(filter(crc_tissue_model_aucs, study == train_model, model_type == "full_model", type == "unmatched")[, "AUC"], filter(crc_tissue_model_aucs, study == train_model, model_type == "select_model", type == "unmatched")[, "AUC"], alternative = "greater", paired = TRUE)$p.value, digits = 3)`). Of the `r length(crc_unmatched_OR_tissue$measure)` taxa with significant ORs, all `r length(crc_unmatched_OR_tissue$measure)` were among the top 10% most important taxa as measured by mean decrease in accuracy, in at least one study. The most important taxa across study within the significant OR taxa only models for stool were *Ruminococcus* and *Clostridium XI* [Figure 5A]. For the singificant OR taxa unmatched tissue models both *Dorea* and *Blautia* were the important based on mean decrease in accuracy [Figure 5B]. These results were surprising because it demonstrated that the ability to classify subjects could be done based on a limited characterization of the communities.

@@

***Generalizability of taxon-based models trained on one dataset to the other datasets:*** 
Considering the good performance of the Random Forest models using taxa with a significant OR and using all of the taxa, we next asked how well the models would perform when given data from a different subject cohort. For instance, if a model was trained using data from the Ahn study, we wanted to know how well it would perform using the data from the Baxter study. We found the models trained using the taxa with a significant OR all had a higher median AUC than the models trained using all of the taxa when tested on the other datasets [Figure 6 & S5]. As might be expected, the difference between the performance of the modelling approaches appeared to vary with the size of the training cohort [Figure 6]. These data suggest that given a sufficient number of subjects with normal colons and carcinomas, Random Forest models trained using a small number of taxa can accurately classify individuals from a different cohort.

@@

**Figure 5: Most Important Members in Significant OR Taxa Carcinoma Models.** A) Common taxa in the top 10 percent for carcinoma Random Forest stool-based models. B) Common taxa in the top 10 percent for carcinoma Random Forest unmatched tissue-based models. Blue represents less important and red represents more important to the Random Forest Model. White means that particular taxa was not in the top 10%.

**Figure 6: Stool Random Forest Genus-Based Model Test AUCs.** A) Test AUCs of adenoma models using all genera across study. B) Test AUCs of carcinoma models using all genera or significant OR taxa only. The black line represents the AUC at 0.5. The red lines represent the median AUC of all test AUCs for a specific study.

@@

**Figure S2: Most Common Taxa Across Carcinoma Full Community Stool Study Models.** A) Common taxa in the top 10 percent for carcinoma Random Forest all taxa-based models. B) Common taxa in the top 10 percent for carcinoma Random Forest all OTU-based models. Blue represents less important and red represents more important to the Random Forest Model. White means that particular taxa was not in the top 10%.

**Figure S3: Most Common Genera Across Full Community Tissue Study Models.** A) Common genera in the top 10 percent for matched carcinoma Random Forest all genera-based models. B) Common genera in the top 10 percent for unmatched carcinoma Random Forest all genera-based models. C) Common genera in the top 10 percent for matched carcinoma Random Forest all OTU-based models. D) Common genera in the top 10 percent for unmatched carcinoma Random Forest all OTU-based models. Blue represents less important and red represents more important to the Random Forest Model. White means that particular taxa was not in the top 10%.

**Figure S4: Tissue Random Forest Model Train AUCs.** A) Adenoma random forest model AUCs between all genera, all OTU, and select model based on significant OR taxa in unmatched and matched tissue. B) Carcinoma random forest model AUCs between all genera, all OTU, and select model based on significant OR taxa in unmatched and matched tissue. The black line represents the median AUC for the respective group. If no values are present in the singificant OR taxa group then there were no significant taxa identified and no model was tested.

**Figure S5: Tissue Random Forest Genus-Based Model Test AUCs.** A) Test AUCs of adenoma models using all genera across study. B) Test AUCs of carcinoma models using all genera for matched tissue studies. C) Test AUCs of carcinoma models using all genera or significant OR taxa only for unmatched tissue studies The black line represents the AUC at 0.5. The red lines represent the median AUC of all test AUCs for a specific study.

### 2018-03-01 submission/manuscript.tex
update with new orders after pats edits

\textbf{\emph{Combined taxa model classifies subjects better than using
individual taxa:}} Instead of attempting to classify subjects based on
individual taxa, next we generated Random Forest models that combined
the individual taxa and evaluated the ability to classify as subject's
disease status. For data from stool samples, the combined model had an
AUC of 0.75, which was significantly higher than any of the AUC values
for the individual taxa (P-value \textless{} 0.033). For the full taxa
models using stool, \emph{Bacteroides} and \emph{Lachnospiraceae} were
the most common taxa in the top 10\% mean decrease in accuracy (MDA)
across studies. Similarly, using data from the unmatched tissue samples,
the combined model had an AUC of 0.77, which was significantly higher
than the AUC values for Blautia and Weissella (P-value \textless{}
0.037). For the full taxa models using unmatched tissue,
\emph{Lachnospiraceae}, \emph{Bacteroidaceae}, and
\emph{Ruminococcaceae} were the most common taxa in the top 10\% mean
decrease in accuracy across studies. Clearly, pooling the information
from the taxa with significant ORs results in a model that outperforms
classifications made using individual taxa.

\textbf{\emph{Performance of models based on taxa relative abundance in
full community are not significantly better than those based on taxa
with significant ORs:}} Next, we asked whether a Random Forest
classification model built using all of the taxa found in the
communities would outperform the models generated using those taxa with
a significant OR. Similar to our inability to identify taxa associated
with a significant OR for the presence of adenomas, the median AUCs to
classify subjects as having normal colons or having adenomas using data
from stool or tissue samples were marginally better than 0.5 for any
study{[}Figure 4A \& S3A{]}. In contrast, the models for classifying
subjects as having normal colons or having carcinomas using data from
stool or tissue samples yielded AUC values meaningfully higher than 0.5
{[}Figure 4B \& S3B-C{]}. When we compared the models based on all of
the taxa in a community to models based on the taxa with significant
ORs, the results were mixed. Using the data from stool samples we found
that although the AUC for 6 of 7 studies increased (mean decrease =
9.53\%), the more expansive models performed worse for 1 of the studies
(decrease = 0.38\%). The overall improvement in performance was
statistically significant (one-tailed paired T-test; P-value = 0.005).
Of the 8 taxa with significant ORs, all 8 were among the top 10\% most
important taxa as measured by mean decrease in accuracy, in at least one
study . Similarly, using the data from unmatched tissue samples we found
that the AUC for 4 out of 4 studies decreased between full versus select
OR models (mean decrease = 19.11\%, one-tailed paired T-test; P-value =
0.03). Of the 3 taxa with significant ORs, all 3 were among the top 10\%
most important taxa as measured by mean decrease in accuracy, in at
least one study. These results were surprising because it demonstrated
that the ability to classify subjects could be done based on a limited
characterization of the communities.

@@

\textbf{\emph{Generalizability of taxon-based models trained on one
dataset to the other datasets:}} Considering the good performance of the
Random Forest models using taxa with a significant OR and using all of
the taxa, we next asked how well the models would perform when given
data from a different subject cohort. For instance, if a model was
trained using data from the Ahn study, we wanted to know how well it
would perform using the data from the Baxter study. We found the models
trained using the taxa with a significant OR all had a higher median AUC
than the models trained using all of the taxa when tested on the other
datasets {[}Figure 5{]}. As might be expected, the difference between
the performance of the modelling approaches appeared to vary with the
size of the training cohort. These data suggest that given a sufficient
number of subjects with normal colons and carcinomas, Random Forest
models trained using a small number of taxa can accurately classify
individuals from a different cohort.

@@

\textbf{Figure 5: Stool Random Forest Genus-Based Model Test AUCs.} A)
Test AUCs of adenoma models using all genera across study. B) Test AUCs
of carcinoma models using all genera or significant OR taxa only. The
black line represents the AUC at 0.5. The red lines represent the median
AUC of all test AUCs for a specific study.

\textbf{Figure 6: Most Common Taxa Across Carcinoma Full Community Stool
Study Models.} A) Common taxa in the top 10 percent for carcinoma Random
Forest all taxa-based models. B) Common taxa in the top 10 percent for
carcinoma Random Forest all OTU-based models.

@@

\textbf{Figure S2: Tissue Random Forest Model Train AUCs.} A) Adenoma
random forest model AUCs between all genera, all OTU, and select model
based on significant OR taxa in unmatched and matched tissue. B)
Carcinoma random forest model AUCs between all genera, all OTU, and
select model based on significant OR taxa in unmatched and matched
tissue. The black line represents the median AUC for the respective
group. If no values are present in the singificant OR taxa group then
there were no significant taxa identified and no model was tested.

\textbf{Figure S3: Tissue Random Forest Genus-Based Model Test AUCs.} A)
Test AUCs of adenoma models using all genera across study. B) Test AUCs
of carcinoma models using all genera for matched tissue studies. C) Test
AUCs of carcinoma models using all genera or significant OR taxa only
for unmatched tissue studies The black line represents the AUC at 0.5.
The red lines represent the median AUC of all test AUCs for a specific
study.

\textbf{Figure S4: Most Common Genera Across Full Community Tissue Study
Models.} A) Common genera in the top 10 percent for matched carcinoma
Random Forest all genera-based models. B) Common genera in the top 10
percent for unmatched carcinoma Random Forest all genera-based models.
C) Common genera in the top 10 percent for matched carcinoma Random
Forest all OTU-based models. D) Common genera in the top 10 percent for
unmatched carcinoma Random Forest all OTU-based models.

>>>>

\textbf{\emph{Combined taxa model classifies subjects better than using
individual taxa:}} Instead of attempting to classify subjects based on
individual taxa, next we generated Random Forest models that combined
the individual taxa and evaluated the ability to classify as subject's
disease status. For data from stool samples, the combined model had an
AUC of 0.75, which was significantly higher than any of the AUC values
for the individual taxa (P-value \textless{} 0.033). For the full taxa
models using stool, \emph{Bacteroides} and \emph{Lachnospiraceae} were
the most common taxa in the top 10\% mean decrease in accuracy (MDA)
across studies {[}Figure S2{]}. Similarly, using data from the unmatched
tissue samples, the combined model had an AUC of 0.77, which was
significantly higher than the AUC values for Blautia and Weissella
(P-value \textless{} 0.037). For the full taxa models using unmatched
tissue, \emph{Lachnospiraceae}, \emph{Bacteroidaceae}, and
\emph{Ruminococcaceae} were the most common taxa in the top 10\% mean
decrease in accuracy across studies {[}Figure S3{]}. Clearly, pooling
the information from the taxa with significant ORs results in a model
that outperforms classifications made using individual taxa.

\textbf{\emph{Performance of models based on taxa relative abundance in
full community are better than those based on taxa with significant
ORs:}} Next, we asked whether a Random Forest classification model built
using all of the taxa found in the communities would outperform the
models generated using those taxa with a significant OR. Similar to our
inability to identify taxa associated with a significant OR for the
presence of adenomas, the median AUCs to classify subjects as having
normal colons or having adenomas using data from stool or tissue samples
were marginally better than 0.5 for any study{[}Figure 4A \& S4A{]}. In
contrast, the models for classifying subjects as having normal colons or
having carcinomas using data from stool or tissue samples yielded AUC
values meaningfully higher than 0.5 {[}Figure 4B \& S4B-C{]}. When we
compared the models based on all of the taxa in a community to models
based on the taxa with significant ORs, the results were mixed. Using
the data from stool samples we found that although the AUC for 6 of 7
studies increased (mean decrease = 9.53\%), the more expansive models
performed worse for 1 of the studies (decrease = 0.38\%). The overall
improvement in performance was statistically significant (one-tailed
paired T-test; P-value = 0.005). Of the 8 taxa with significant ORs, all
8 were among the top 10\% most important taxa as measured by mean
decrease in accuracy, in at least one study . Similarly, using the data
from unmatched tissue samples we found that the AUC for 4 out of 4
studies decreased between full versus select OR models (mean decrease =
19.11\%, one-tailed paired T-test; P-value = 0.03). Of the 3 taxa with
significant ORs, all 3 were among the top 10\% most important taxa as
measured by mean decrease in accuracy, in at least one study. The most
important taxa across study within the significant OR taxa only models
for stool were \emph{Ruminococcus} and \emph{Clostridium XI} {[}Figure
5A{]}. For the singificant OR taxa unmatched tissue models both
\emph{Dorea} and \emph{Blautia} were the important based on mean
decrease in accuracy {[}Figure 5B{]}. These results were surprising
because it demonstrated that the ability to classify subjects could be
done based on a limited characterization of the communities.

@@

\textbf{\emph{Generalizability of taxon-based models trained on one
dataset to the other datasets:}} Considering the good performance of the
Random Forest models using taxa with a significant OR and using all of
the taxa, we next asked how well the models would perform when given
data from a different subject cohort. For instance, if a model was
trained using data from the Ahn study, we wanted to know how well it
would perform using the data from the Baxter study. We found the models
trained using the taxa with a significant OR all had a higher median AUC
than the models trained using all of the taxa when tested on the other
datasets {[}Figure 6 \& S5{]}. As might be expected, the difference
between the performance of the modelling approaches appeared to vary
with the size of the training cohort {[}Figure 6{]}. These data suggest
that given a sufficient number of subjects with normal colons and
carcinomas, Random Forest models trained using a small number of taxa
can accurately classify individuals from a different cohort.

@@

\textbf{Figure 5: Most Important Members in Significant OR Taxa
Carcinoma Models.} A) Common taxa in the top 10 percent for carcinoma
Random Forest stool-based models. B) Common taxa in the top 10 percent
for carcinoma Random Forest unmatched tissue-based models. Blue
represents less important and red represents more important to the
Random Forest Model. White means that particular taxa was not in the top
10\%.

\textbf{Figure 6: Stool Random Forest Genus-Based Model Test AUCs.} A)
Test AUCs of adenoma models using all genera across study. B) Test AUCs
of carcinoma models using all genera or significant OR taxa only. The
black line represents the AUC at 0.5. The red lines represent the median
AUC of all test AUCs for a specific study.

@@

\textbf{Figure S2: Most Common Taxa Across Carcinoma Full Community
Stool Study Models.} A) Common taxa in the top 10 percent for carcinoma
Random Forest all taxa-based models. B) Common taxa in the top 10
percent for carcinoma Random Forest all OTU-based models. Blue
represents less important and red represents more important to the
Random Forest Model. White means that particular taxa was not in the top
10\%.

\textbf{Figure S3: Most Common Genera Across Full Community Tissue Study
Models.} A) Common genera in the top 10 percent for matched carcinoma
Random Forest all genera-based models. B) Common genera in the top 10
percent for unmatched carcinoma Random Forest all genera-based models.
C) Common genera in the top 10 percent for matched carcinoma Random
Forest all OTU-based models. D) Common genera in the top 10 percent for
unmatched carcinoma Random Forest all OTU-based models. Blue represents
less important and red represents more important to the Random Forest
Model. White means that particular taxa was not in the top 10\%.

\textbf{Figure S4: Tissue Random Forest Model Train AUCs.} A) Adenoma
random forest model AUCs between all genera, all OTU, and select model
based on significant OR taxa in unmatched and matched tissue. B)
Carcinoma random forest model AUCs between all genera, all OTU, and
select model based on significant OR taxa in unmatched and matched
tissue. The black line represents the median AUC for the respective
group. If no values are present in the singificant OR taxa group then
there were no significant taxa identified and no model was tested.

\textbf{Figure S5: Tissue Random Forest Genus-Based Model Test AUCs.} A)
Test AUCs of adenoma models using all genera across study. B) Test AUCs
of carcinoma models using all genera for matched tissue studies. C) Test
AUCs of carcinoma models using all genera or significant OR taxa only
for unmatched tissue studies The black line represents the AUC at 0.5.
The red lines represent the median AUC of all test AUCs for a specific
study.
